Asymptotic and Numerical Solution for Electro-Kinetic Flow Around a Spherical Particle

### 2012-09-11 notes/work/ijnam.tex
work on LaTeX

Electrokinetic theory describes the dynamics of charged particles
in ionic fluids \cite{masliyah2005book,kirby2010book}.
When a particle acquires surface charge, a layer
of ions of opposite charge is attracted to the surface via    
electric forces, creating a double-layer structure around the
particle (see Figure \ref{fig:EDL}). This structure, called
``Debye layer'', electrically screens the surface charge, thus
creating a potential difference between the particle and the outer
layer of the fluid bulk.
In cases where the layer width is much smaller than the particle
size, an analytical asymptotic solution for the Debye layer
dynamics can be found \cite{yariv2010asymptotic}.
\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=0.33\textwidth]
            {figs/debye.eps}
        \caption{Schematic structure of the double layer:
        (1) particle surface, (2) Debye layer, (3) fluid bulk.
        If the particle surface has positive (blue) surface charge,
        it attracts negative (red) ions from the fluid making the
        Debye layer negatively charged (as opposed to the rest of
        the fluid bulk, which is electrically neutral).
        The zeta potential is defined as the voltage drop across 
        the Debye layer (2).}
        \label{fig:EDL}
    \end{center}
\end{figure}
% \begin{tikzpicture}\end{tikzpicture}

@@

The goal of this work is to develop and implement an iterative numerical
solver for nonlinear electrokinetic problems, and apply it to the study
of systems that have no closed-form solutions, such as an ion-exchanger migration
\cite{yariv2010migration} and the electrophoresis problem \cite{schnitzer2012surface}.
This solver can be used to gain insight into the chemical and physical behavior 
in far more general regimes than are currently well-understood, 
and also to possibly lead to further analytical developments, based on new scaling behavior
discovered numerically. We may also target interesting physical
phenomena that have been observed experimentally.
One such phenomenon is nonlinear electrokinetic vortex flow, 
which is useful for microfluidic mixing applications \cite{wang2004mix, ben2002vortex},

@@

After nondimensionalization, the Nernst-Plank, Stokes and Poisson equations are:
\begin{eqnarray} \label{eq:nernst}
\bj_\pm &=& 
-\bnabla c_\pm + \alpha \bv c_\pm \mp c_\pm \bnabla \varphi,
\end{eqnarray}
\begin{eqnarray} \label{eq:stokes}
\bnabla \cdot \bv = 0, & -\bnabla p + \bLaplacian \bv + \Laplacian \varphi \bnabla \varphi = 0, 
\end{eqnarray}
\begin{eqnarray} \label{poisson}
c_+ - c_- &=& -2\delta^2 \Laplacian \varphi,
\end{eqnarray}
where $\delta = {\delta^*}/{a^*}$ is the dimensionless Debye thickness 
($\delta^* = \sqrt{{\eps^* \varphi^*}/{2 c^* e^* z}}$) and 
$\alpha = {v^* a^*}/{D^*} = {\eps^* (\varphi^*)^2}/{\mu^* D^*}$ 
is the dimensionless Peclet number.
The force on the particle is denoted by $\bF$ normalized by $\mu^* v^* a^* = \eps^* (\varphi^*)^2$.
The stress tensor is denoted by $\tT$ and normalized by $p^* = \mu^* v^* / a^*$.
The electric field is denoted by $\bE$ and normalized by $E^* = \varphi^* / a^*$.
The surface charge is denoted by $\sigma$ and normalized by $\sigma^* = \eps^* \varphi^* / \delta^*$.

The ionic fluxes are given by ion diffusion, electrostatic forces and ion advection by the fluid
at Eq. \ref{eq:nernst}. Due to conservation of ions, the fluxes are divergence-free:
\begin{eqnarray*}
\bnabla \cdot \bj_\pm &=& 0.
\end{eqnarray*}
These equation may be re-written by using salt $c = (c_+ + c_-)/2$ and charge $q = (c_+ - c_-)/2$.
Thus, salt and charge fluxes are defined by:
\begin{eqnarray*}
  \bj = \frac{\bj_+ + \bj_-}{2} = -\bnabla c - q \bnabla \varphi + \alpha \bv c &,& 
  \bi = \frac{\bj_+ - \bj_-}{2} = -\bnabla q - c \bnabla \varphi + \alpha \bv q,
\end{eqnarray*}
and are divergence-free as well:
\begin{eqnarray*}
\bnabla \cdot \bj = 0 &,& 
\bnabla \cdot \bi = 0. 
\end{eqnarray*}
In addition, the charge is governed by the Poisson equation:
\begin{eqnarray*}
q &=& -\delta^2 \Laplacian \varphi.
\end{eqnarray*}

@@

We employ the reference frame of the particle.
Far away from it, we have a uniform applied electrostatic field, uniform flow,
and ambient ionic concentrations:
\begin{eqnarray*}
\bv \rightarrow -\cU \ui, &
\bnabla \varphi \rightarrow -\beta\ui, &
 c_\pm \rightarrow 1.
\end{eqnarray*}

We consider a spherical particle.
The chemical properties of the particle's surface $\mathcal{S}$ (defined at $r=1$ 
with normal $\bn = \brhat$), determine the boundary conditions on $\mathcal{S}$.
We consider the following scenarios:
\subsubsection{Ion-exchanger (\cite{yariv2010migration})}
We assume high conductance of the particle (yielding a uniform electric potential),
zero slip, anion impermeability, cation selectivity 
(with fast Butler-Volmer\cite{bard2000book} kinetics
$\bn \cdot \bj_+ = k (1 - c_+/\gamma)$, where $k \rightarrow \infty$):
\begin{eqnarray*}
\begin{array}{cccc}
\varphi = \cV, &
\bv = \bzero, &
\bn \cdot \bj_- = 0, &
c_+ = \gamma,
\end{array}
\end{eqnarray*}
where $\gamma$ is the ratio between the cationic concentration on the surface and the cationic
concentration in the fluid bulk.

\subsubsection{Inert particle (\cite{schnitzer2012surface})}
We assume constant surface charge
zero slip and ion impermeability:
\begin{eqnarray*}
\begin{array}{cccc}
\bn \cdot (\bnabla \varphi - \chi \bnabla \varphi_s) = \delta^{-1} \sigma, &
\bv = \bzero, &
\bn \cdot \bj_\pm = 0.
\end{array}
\end{eqnarray*}
where $\chi$ is the ratio between the dielectric constants of the 
particle and the fluid.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Separation of scales}
When $\delta \ll 1$, 
most of the fluid remains electroneutral ($q \approx 0$, thus $c_+ \approx c_-$), 
except a thin boundary layer of width $O(\delta)$, 
known as the ``Debye layer'', surrounding the particle.
This boundary layer makes it very hard to construct 
a direct numerical solver of the electro-kinetic problem.

Due to the scale disparity at the Debye layer, 
the PDE system can be rewritten in the ``inner'' region 
as an ODE system and can be integrated and solved analytically,
yielding an ``effective'' boundary conditions for the ``outer'' bulk region equations.

\subsubsection  {Bulk-scale equations}
The Nernst-Planck equations above for the fluid bulk (outside
the boundary layer), using bulk variables denoted by uppercase letters, read:
\begin{eqnarray*}
\begin{array}{cccc}
  Q = 0, &
  C = C_+ = C_-, &
\bJ = -\bnabla C + \alpha \bV C, &
\bI = -C \bnabla \varPhi.
\end{array}\end{eqnarray*}
Because the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following 
salt and charge conservation equations:
\begin{eqnarray*} 
0 &=& \bnabla \cdot \bJ = \Laplacian C - \alpha \bV \cdot \bnabla C, \\
0 &=& \bnabla \cdot \bI = \bnabla \cdot \pars{ C \bnabla \varPhi }.
\end{eqnarray*}
The macroscale Stokes flow equations have the same form as the microscale equations:
\begin{eqnarray*} 
0 &=& \bnabla \cdot \bV, \\   
\bzero &=& \Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi.
\end{eqnarray*}

\subsection{Bulk-scale effective boundary conditions}
By integrating the equations across the Debye layer,
the effective boundary conditions at the particle surface are:
\subsubsection{Ion-exchanger} See \cite{yariv2010migration} for full derivation:
\begin{eqnarray*} 
0 &=& \varPhi + \log C, \\ 
0 &=& \deriv{}{n} \pars{\varPhi - \log C}, \\ 
\bV &=& 
\zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C,
\end{eqnarray*}
where $\zeta = \cV - \varPhi = \log (C / \gamma)$ 
is the zeta potential (the voltage drop across the Debye layer).

\subsubsection{Highly charged inert particle}
See \cite{schnitzer2012surface} for full derivation:
\begin{eqnarray*} 
0 &=& {\deriv{C}{n} - \text{Du} \nabla^2_s \pars{\varPhi - \log C}} \\ 
0 &=& {\deriv{C}{n} + C \deriv{\varPhi}{n}}, \\ 
\bV &=& \zeta \cdot \bnabla_\mathcal{S} \pars{\varPhi - \log C} +
4 \log 2 \cdot \bnabla_\mathcal{S} \log C,
\end{eqnarray*}
where $\zeta = \bar{\zeta} - \log C$ 
is the zeta potential (the voltage drop across the Debye layer)
and $\bar{\zeta} = 2 \log \sigma$.

\subsection{Steady state}
The stress tensor is composed of Newtonian and Maxwel stresses:
\begin{eqnarray*} 
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi) \tI, \\
 \bnabla \cdot \tT &=& \bzero.
\end{eqnarray*}
In steady state (constant particle drift velocity $\cU$), the total force acting on the particle 
(computed by the following surface integral) must vanish:
\begin{eqnarray*}
 \bF = \oint_\mathcal{S} \tT \cdot \bnhat dA = \bzero.
\end{eqnarray*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithm Description} \label{sec:algorithm}
The numerical solver first requires constructing a discrete 
approximation for the nonlinear PDE system,
for a given electric field $\beta$ and a velocity $\cU$. 
The resulting nonlinear discrete system is
solved numerically using Newton's Method, so the total force on the
particle can be computed. 
We seek the steady-state velocity $\cU$, 
which corresponds to zero total force,
\begin{eqnarray*} 
F(\beta, \cU) = 0,
\end{eqnarray*}
which we solve by the standard secant method.

\subsection{Discretization}
\subsubsection{Spherical Coordinates}
Because the system is axisymmetric, we employ
spherical coordinates $(r,\theta,\phi)$.
The scalar gradient and the divergence can be written as:
\begin{eqnarray*}
\bnabla f &=& \deriv{f}{r}\brhat + \frac{1}{r}\deriv{f}{\theta}\bthetahat, \\
\bnabla \cdot \bF &=& \frac{1}{r^2}\deriv{}{r}\pars{F_r r^2 } + 
               \frac{1}{r \sin\theta}\deriv{}{\theta}\pars{F_\theta \sin\theta}.
\end{eqnarray*}
The scalar Laplacian can be written as:
\begin{eqnarray*}
\Laplacian f = \bnabla \cdot (\bnabla f)&=& 
 \frac{1}{r^2}\deriv{}{r}\pars{r^2 \deriv{f}{r}} + 
 \frac{1}{r^2 \sin\theta}\deriv{}{\theta}\pars{\sin\theta \deriv{f}{\theta}}.
\end{eqnarray*}
The vector gradient can be written as:
\begin{eqnarray*}
\bnabla \bF &=& \deriv{F_r}{r} \brhat \brhat + \deriv{F_\theta}{r} \brhat \bthetahat + 
\frac{1}{r}\pars{\deriv{F_r}{\theta} - F_\theta} \bthetahat \brhat + 
\frac{1}{r}\pars{\deriv{F_\theta}{\theta} + F_r} \bthetahat \bthetahat.
\end{eqnarray*}
The vector Laplacian can be written as:
\begin{eqnarray*}
\bLaplacian \bF &=& 
\left(\Laplacian F_r - \frac{2F_r}{r^2} - 
\frac{2}{r^2 \sin\theta} \deriv{\left(F_\theta \sin\theta \right)}{\theta}\right)\brhat
+ \left(\Laplacian F_\theta - \frac{F_\theta}{r^2 \sin^2\theta} + 
\frac{2}{r^2}\deriv{F_r}{\theta}\right) \bthetahat.
\end{eqnarray*}

@@

\subsubsection{Computational Grid}
A regular grid of size $n_r \times n_\theta$ is defined, with:
\begin{eqnarray*}
(r_i,\theta_j) \in [1, \infty) \times [0,\pi], & 
r_i = (1+\Delta_r)^i, &
 \theta_j = \Delta_\theta \cdot j,
\end{eqnarray*}
where a logarithmic grid spacing has been chosen for $r$ and a uniform grid for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{n_r}$ should be much 
larger than $r_0$. Thus:
\begin{eqnarray*}
\Delta_r = \pars{R_{max}} ^ \frac{1}{n_r} - 1 ,& \Delta_\theta = \frac{\pi}{n_\theta}.
\end{eqnarray*}

This grid induces a disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi] = \bigcup_{ij}\Omega_{ij}$ into cells.
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by:
\begin{eqnarray*}
\Omega_{ij} = [r_{i-1}, r_{i}] \times [\theta_{j-1}, \theta_{j}], &
\bar{r}_i = (r_{i-1} + r_{i})/2, &
\bar{\theta}_j = (\theta_{j-1} + \theta_{j})/2.
\end{eqnarray*}

@@

\subsubsection{Operator Discretization}
A finite-volume method with linear interpolation is used for flux discretization.

@@

Ion flux ($\bI = - C \bnabla \varPhi$) is discretized on grid cell boundaries:
\begin{eqnarray*}
I^h_r = -\cI_r(C^h) \cdot \cD_r(\varPhi^h), &
I^h_\theta = -\cI_\theta(C^h) \cdot \frac{\cD_\theta(\varPhi^h)}{r}.
\end{eqnarray*}

Salt flux ($\bJ = -\bnabla C + \alpha \bV C$) is discretized on grid cell boundaries 
(using upwind scheme $\mathcal{U}$ for numerical stability at large cell Peclet number):
\begin{eqnarray*}
J^h_r = -\cD_r(C^h) + \alpha V^h_r \cdot \mathcal{U}^{\bV^h}_r (C^h), &
J^h_\theta = -\frac{\cD_\theta(C^h)}{r} + \alpha V^h_\theta \cdot \mathcal{U}^{\bV^h}_\theta (C^h), 
\\
 \mathcal{U}^{\bV^h}_r(C^h)[i,j] = C^h\left[i-\frac{\sign(V^h_r)}{2}, j\right], &
 \mathcal{U}^{\bV^h}_\theta(C^h)[i,j] = C^h\left[i, j-\frac{\sign(V^h_\theta)}{2}\right]. 
\end{eqnarray*}

@@

Force components ($\bF = -\bnabla P + \bLaplacian \bV + \bnabla \varPhi \Laplacian \varPhi$) 
are discretized on the staggered velocity grid, where
linear interpolation is used for the Coulomb force and for vector Laplacian components.
\begin{eqnarray*}
F^h_r &=& -\cD_r(P^h) 
          + \cL(V^h_r) - \frac{2}{r^2} V^h_r 
		  - \frac{2}{r^2 \sin\theta} \cI_r(\cD_\theta (V^h_\theta \sin\theta))
          + \cD_r(\varPhi^h) \cdot \cI_r(\cL(\varPhi^h)), \\
F^h_\theta &=& -\frac{\cD_\theta(P^h)}{r} 
		  + \cL(V^h_\theta) - \frac{F^h_\theta}{r^2 \sin^2\theta} 
		  + \frac{2}{r^2} \cI_\theta(\cD_\theta(F^h_r))
		  + \frac{\cD_\theta(\varPhi^h)}{r} \cdot \cI_\theta(\cL(\varPhi^h)), \\
\cL(f^h) &=& \frac{1}{r^2}\cD_r\pars{\cD_r(f^h) r^2} + 
\frac{1}{r^2 \sin\theta} \cD_\theta\pars{\cD_\theta(f^h) \cdot \sin\theta}.
\end{eqnarray*}

\subsubsection{Boundary Conditions}
In order to discretize boundary condtions for $\varPhi$, $C$ and $\bV$, 
``ghost'' points are employed. 
The grid is extended to include points outside the domain interior,
and the variables at the ``ghost'' points are set to satisfy 
the discretized boundary conditions, following the analysis below.
Note that $P$ has no boundary conditions -- so the pressure variable is defined 
only in the domain interior.

>>>>

Electrokinetic theory describes the dynamics of charged particles
in ionic fluids \cite{masliyah2005book,kirby2010book}.
When a particle acquires surface charge, a layer
of ions of opposite charge is attracted to the surface via    
electric forces, creating a double-layer structure around the
particle (see Figure \ref{fig:EDL}). This structure, called
``Debye layer'', electrically screens the surface charge, thus
creating a potential difference between the particle and the outer
layer of the fluid bulk.
In cases where the layer width is much smaller than the particle
size, an analytical asymptotic solution for the Debye layer
dynamics can be found \cite{yariv2010asymptotic}.
\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=0.35\textwidth]
            {figs/debye.eps}
        \caption{Schematic structure of the double layer:
        (1) particle surface, (2) Debye layer, (3) fluid bulk.
        If the particle surface has positive (blue) surface charge,
        it attracts negative (red) ions from the fluid making the
        Debye layer negatively charged (as opposed to the rest of
        the fluid bulk, which is electrically neutral).
        The zeta potential is defined as the voltage drop across 
        the Debye layer (2).}
        \label{fig:EDL}
    \end{center}
\end{figure}
% \begin{tikzpicture}\end{tikzpicture}

@@

The goal of this work is to develop and implement a software toolbox for the generation
of iterative numerical solvers for nonlinear electrokinetic problems, and apply it to the study
of systems that have no closed-form solutions, such as an ion-exchanger migration
\cite{yariv2010migration} and the electrophoresis problem \cite{schnitzer2012surface}.
This solver can be used to gain insight into the chemical and physical behavior 
in far more general regimes than are currently well-understood, 
and also to possibly lead to further analytical developments, based on new scaling behavior
discovered numerically. We may also target interesting physical
phenomena that have been observed experimentally.
One such phenomenon is nonlinear electrokinetic vortex flow, 
which is useful for microfluidic mixing applications \cite{wang2004mix, ben2002vortex},

@@

After nondimensionalization, the Nernst-Plank, Stokes and Poisson equations are:
\begin{eqnarray} \label{eq:nernst}
\bj_\pm &=& 
-\bnabla c_\pm + \alpha \bv c_\pm \mp c_\pm \bnabla \varphi,
\end{eqnarray}
\begin{eqnarray} \label{eq:stokes}
\bnabla \cdot \bv = 0, & -\bnabla p + \bLaplacian \bv + \Laplacian \varphi \bnabla \varphi = 0, 
\end{eqnarray}
\begin{eqnarray} \label{eq:poisson}
c_+ - c_- &=& -2\delta^2 \Laplacian \varphi,
\end{eqnarray}
where $\delta = {\delta^*}/{a^*}$ is the dimensionless Debye thickness 
($\delta^* = \sqrt{{\eps^* \varphi^*}/{2 c^* e^* z}}$) and 
$\alpha = {v^* a^*}/{D^*} = {\eps^* (\varphi^*)^2}/{\mu^* D^*}$ 
is the dimensionless Peclet number.
The force on the particle is denoted by $\bF$ normalized by $\mu^* v^* a^* = \eps^* (\varphi^*)^2$.
The stress tensor is denoted by $\tT$ and normalized by $p^* = \mu^* v^* / a^*$.
The electric field is denoted by $\bE$ and normalized by $E^* = \varphi^* / a^*$.
The surface charge is denoted by $\sigma$ and normalized by $\sigma^* = \eps^* \varphi^* / \delta^*$.

The ionic fluxes are given by ion diffusion, electrostatic forces and ion advection by the fluid
at Eq. \eqref{eq:nernst}. Due to conservation of ions, the fluxes are divergence-free:
\begin{eqnarray*}
\bnabla \cdot \bj_\pm &=& 0.
\end{eqnarray*}
These equations may be re-written by using salt $c = (c_+ + c_-)/2$ and charge $q = (c_+ - c_-)/2$.
Thus, salt and charge fluxes are defined by:
\begin{eqnarray} \label{eq:fluxes}
  \bj = \frac{\bj_+ + \bj_-}{2} = -\bnabla c - q \bnabla \varphi + \alpha \bv c &,& 
  \bi = \frac{\bj_+ - \bj_-}{2} = -\bnabla q - c \bnabla \varphi + \alpha \bv q,
\end{eqnarray}
and are divergence-free as well:
\begin{eqnarray} \label{eq:zero_flux}
\bnabla \cdot \bj = 0 &,& 
\bnabla \cdot \bi = 0. 
\end{eqnarray}
In addition, the charge concentration is governed by the Poisson equation \eqref{eq:poisson}
\begin{eqnarray}
q &=& -\delta^2 \Laplacian \varphi.
\end{eqnarray}

@@

We employ the reference frame of the particle.
Far away from it, we have a uniform applied electrostatic field, uniform flow,
and ambient ionic concentrations:
\begin{eqnarray} \label{eq:bndcond_inf}
\bv \rightarrow -\cU \ui, &
\bnabla \varphi \rightarrow -\beta\ui, &
 c_\pm \rightarrow 1.
\end{eqnarray}

We consider a spherical particle.
The chemical properties of the particle's surface $\mathcal{S}$ (defined at $r=1$ 
with normal $\bn = \brhat$), determine the boundary conditions on $\mathcal{S}$.
We consider the following scenarios:
\subsubsection{Ion-exchanger (\cite{yariv2010migration})}
We assume high conductance of the particle (yielding a uniform electric potential),
zero slip, anion impermeability, cation selectivity 
(with fast Butler-Volmer\cite{bard2000book} kinetics
$\bn \cdot \bj_+ = k (1 - c_+/\gamma)$, where $k \rightarrow \infty$):
\begin{eqnarray}
\begin{array}{cccc}
\varphi = \cV, &
\bv = \bzero, &
\bn \cdot \bj_- = 0, &
c_+ = \gamma,
\end{array}
\end{eqnarray}
where $\gamma$ is the ratio between the cationic concentration on the surface and the cationic
concentration in the fluid bulk.

\subsubsection{Inert particle (\cite{schnitzer2012surface})}
We assume constant surface charge
zero slip and ion impermeability:
\begin{eqnarray}
\begin{array}{cccc}
\bn \cdot (\bnabla \varphi - \chi \bnabla \varphi_s) = \delta^{-1} \sigma, &
\bv = \bzero, &
\bn \cdot \bj_\pm = 0.
\end{array}
\end{eqnarray}
where $\chi$ is the ratio between the dielectric constants of the 
particle and the fluid.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Separation of scales}
When $\delta \ll 1$, 
most of the fluid remains electroneutral ($q \approx 0$, thus $c_+ \approx c_-$), 
except a thin boundary layer of width $O(\delta)$, 
known as the ``Debye layer'', surrounding the particle.
This boundary layer makes it very hard to construct 
a direct numerical solver for the full electro-kinetic problem.

Due to the scale disparity at the Debye layer, 
the PDE system can be rewritten in the ``inner'' region 
as an ODE system using the asymptotic limit of $\delta \rightarrow 0$ 
and can be integrated and solved analytically,
yielding an ``effective'' boundary conditions for the ``outer'' bulk region equations near
the particle surface.

\subsubsection  {Bulk-scale equations}
The Nernst-Planck equations above \eqref{eq:fluxes} for the fluid bulk (outside
the boundary layer), using bulk variables denoted by uppercase letters, read:
\begin{eqnarray} \begin{array}{cccc} \label{eq:bulk_flux}
  Q = 0, &
  C = C_+ = C_-, &
\bJ = -\bnabla C + \alpha \bV C, &
\bI = -C \bnabla \varPhi.
\end{array}\end{eqnarray}
Because the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following 
salt and charge conservation equations -- the salt and charge fluxes above 
\eqref{eq:bulk_flux} are divergence-free
\eqref{eq:zero_flux}:
\begin{eqnarray} 
\label{eq:salt_charge}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0, &
\bnabla \cdot \pars{ C \bnabla \varPhi } = 0.
\end{eqnarray}
The macroscale Stokes flow equations have the same form as before \eqref{eq:stokes}:
\begin{eqnarray} 
\bnabla \cdot \bV = 0, &   
\Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi = \bzero.
\end{eqnarray}

\subsection{Bulk-scale effective boundary conditions}
By integrating the equations across the Debye layer,
the effective boundary conditions at the particle surface are:
\subsubsection{Ion-exchanger} See \cite{yariv2010migration} for full derivation:
\begin{eqnarray} \label{eq:ionex_bnd}
0 = \varPhi + \log C, &
0 = \deriv{}{n} \pars{\varPhi - \log C}, &
\bV = 
\zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C,
\end{eqnarray}
where $\zeta = \cV - \varPhi = \log (C / \gamma)$ 
is the zeta potential.

\subsubsection{Highly charged inert particle}
See \cite{schnitzer2012surface} for full derivation:
\begin{eqnarray} \label{eq:ephor_bnd}
0 = {\deriv{C}{n} - \text{Du} \nabla^2_s \pars{\varPhi - \log C}}, & 
0 = {\deriv{C}{n} + C \deriv{\varPhi}{n}}, &
\bV = \zeta \cdot \bnabla_\mathcal{S} \pars{\varPhi - \log C} +
4 \log 2 \cdot \bnabla_\mathcal{S} \log C,
\end{eqnarray}
where $\zeta = \bar{\zeta} - \log C$ 
is the zeta potential and $\bar{\zeta} = 2 \log \sigma$.

\subsection{Steady state}
The stress tensor in the fluid is composed of Newtonian and Maxwel stresses:
\begin{eqnarray} \label{eq:tensor}
\tT &=& \bnabla \bV + (\bnabla \bV)^\dagger - P \tI
+ \bnabla \varPhi \bnabla \varPhi - \frac{1}{2} (\bnabla \varPhi \cdot \bnabla \varPhi). 
\end{eqnarray}
In steady state (constant particle drift velocity $\cU$), the total force acting on the particle 
(computed by the following surface integral) must vanish:
\begin{eqnarray} \label{eq:zero_force}
 \bF = \oint_\mathcal{S} \tT \cdot \bnhat dA = \bzero.
\end{eqnarray}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithm Description} \label{sec:algorithm}
The numerical solver first requires constructing a discrete 
approximation for the nonlinear PDE system,
for a given electric field $\beta$ and a velocity $\cU$. 
The resulting nonlinear discrete system is
solved numerically using Newton's Method, so the total force on the
particle can be computed. 
We seek the steady-state velocity $\cU$, 
which corresponds to zero total force, $F(\beta, \cU) = 0$,
which we solve by the standard secant method.

\subsection{Spherical Coordinates}
Because the system is axisymmetric, we employ
spherical coordinates $(r,\theta,\phi)$.
The scalar gradient and the divergence can be written as:
\begin{eqnarray*}
\bnabla f &=& \deriv{f}{r}\brhat + \frac{1}{r}\deriv{f}{\theta}\bthetahat, \\
\bnabla \cdot \bF &=& \frac{1}{r^2}\deriv{}{r}\pars{F_r r^2 } + 
               \frac{1}{r \sin\theta}\deriv{}{\theta}\pars{F_\theta \sin\theta}.
\end{eqnarray*}
The scalar Laplacian can be written as:
\begin{eqnarray*}
\Laplacian f = \bnabla \cdot (\bnabla f)&=& 
 \frac{1}{r^2}\deriv{}{r}\pars{r^2 \deriv{f}{r}} + 
 \frac{1}{r^2 \sin\theta}\deriv{}{\theta}\pars{\sin\theta \deriv{f}{\theta}}.
\end{eqnarray*}
The vector gradient can be written as:
\begin{eqnarray*}
\bnabla \bF &=& \deriv{F_r}{r} \brhat \brhat + \deriv{F_\theta}{r} \brhat \bthetahat + 
\frac{1}{r}\pars{\deriv{F_r}{\theta} - F_\theta} \bthetahat \brhat + 
\frac{1}{r}\pars{\deriv{F_\theta}{\theta} + F_r} \bthetahat \bthetahat.
\end{eqnarray*}
The vector Laplacian can be written as:
\begin{eqnarray*}
\bLaplacian \bF &=& 
\left(\Laplacian F_r - \frac{2F_r}{r^2} - 
\frac{2}{r^2 \sin\theta} \deriv{\left(F_\theta \sin\theta \right)}{\theta}\right)\brhat
+ \left(\Laplacian F_\theta - \frac{F_\theta}{r^2 \sin^2\theta} + 
\frac{2}{r^2}\deriv{F_r}{\theta}\right) \bthetahat.
\end{eqnarray*}

@@

\subsection{Computational Grid}
A regular grid of size $n_r \times n_\theta$ is defined, with:
\begin{eqnarray*}
(r_i,\theta_j) \in [1, \infty) \times [0,\pi], & 
r_i = (1+\Delta_r)^i, &
 \theta_j = \Delta_\theta \cdot j,
\end{eqnarray*}
where a logarithmic grid spacing has been chosen for $r$ and a uniform grid for $\theta$.
Note that $r_0 = 1$ and $R_{max} = (1+\Delta_r)^{n_r}$, where
\begin{eqnarray*}
\Delta_r = \pars{R_{max}} ^ \frac{1}{n_r} - 1 ,& \Delta_\theta = \frac{\pi}{n_\theta}.
\end{eqnarray*}
Because the grid is finite, the choice 
$R_{max}$ must be taken large enough so as to have a negligible effect
on the solution.
This grid induces a disjoint subdivision of the domain 
$\Omega = [1, R_{max}] \times [0,\pi] = \bigcup_{ij}\Omega_{ij}$ into cells.
A specific cell $\Omega_{ij}$ and its center $(\bar{r}_i, \bar{\theta}_j)$ are defined by:
\begin{eqnarray*}
\Omega_{ij} = [r_{i-1}, r_{i}] \times [\theta_{j-1}, \theta_{j}], &
\bar{r}_i = (r_{i-1} + r_{i})/2, &
\bar{\theta}_j = (\theta_{j-1} + \theta_{j})/2.
\end{eqnarray*}

@@

\subsection{Operator Discretization} 
A finite-volume method with linear interpolation is used for flux discretization.

@@

\subsection{System Equations} \label{sec:disc_equations}
Charge and salt fluxes \eqref{eq:bulk_flux} are discretized on grid cell boundaries 
(using upwind scheme $\mathcal{U}$ for numerical stability at large cell Peclet number):
\begin{eqnarray} \nonumber
I^h_r = -\cI_r(C^h) \cdot \cD_r(\varPhi^h), &
I^h_\theta = -\cI_\theta(C^h) \cdot \frac{\cD_\theta(\varPhi^h)}{r}, \\
J^h_r = -\cD_r(C^h) + \alpha V^h_r \cdot \mathcal{U}^{\bV^h}_r (C^h), &
J^h_\theta = -\frac{\cD_\theta(C^h)}{r} + \alpha V^h_\theta \cdot \mathcal{U}^{\bV^h}_\theta (C^h), 
\\ \nonumber
 \mathcal{U}^{\bV^h}_r(C^h)[i,j] = C^h\left[i-\frac{\sign(V^h_r)}{2}, j\right], &
 \mathcal{U}^{\bV^h}_\theta(C^h)[i,j] = C^h\left[i, j-\frac{\sign(V^h_\theta)}{2}\right]. 
\end{eqnarray}

@@

Force components ($\bF = -\bnabla P + \bLaplacian \bV + \bnabla \varPhi \Laplacian \varPhi$) 
are discretized on the staggered velocity grid, where
linear interpolation is used for the Coulomb force and for vector Laplacian components.
\begin{eqnarray}
\nonumber
F^h_r &=& -\cD_r(P^h) 
          + \cL(V^h_r) - \frac{2}{r^2} V^h_r 
		  - \frac{2}{r^2 \sin\theta} \cI_r(\cD_\theta (V^h_\theta \sin\theta))
          + \cD_r(\varPhi^h) \cdot \cI_r(\cL(\varPhi^h)), \\
F^h_\theta &=& -\frac{\cD_\theta(P^h)}{r} 
		  + \cL(V^h_\theta) - \frac{F^h_\theta}{r^2 \sin^2\theta} 
		  + \frac{2}{r^2} \cI_\theta(\cD_\theta(F^h_r))
		  + \frac{\cD_\theta(\varPhi^h)}{r} \cdot \cI_\theta(\cL(\varPhi^h)), \\
\nonumber
\cL(f^h) &=& \frac{1}{r^2}\cD_r\pars{\cD_r(f^h) r^2} + 
\frac{1}{r^2 \sin\theta} \cD_\theta\pars{\cD_\theta(f^h) \cdot \sin\theta}.
\end{eqnarray}

\subsection{Boundary Conditions} \label{sec:disc_boundary}
In order to discretize boundary condtions for $\varPhi$, $C$ and $\bV$, 
``ghost'' points are employed. 
The grid is extended to include points outside the domain interior,
and the variables at the ``ghost'' points are set to satisfy 
the discretized boundary conditions, following the analysis below.
Note that $P$ has no boundary conditions -- so the pressure variable is defined 
only in the domain interior.

Far away from the particle boundary \eqref{eq:bndcond_inf}, we have:
\begin{eqnarray*}
\begin{array}{cccc}
\cD_r \pars{\varPhi^h} = -\beta \cos\bar{\theta}, &
\cI_r \pars{C^h} = 1 &
V_r^h = -\cU \cos\theta, &
V_\theta^h = \cU \sin\theta.
\end{array}\end{eqnarray*}

For $\theta = 0$ and $\theta = \pi$, the ``ghost'' points are defined by axial 
symmetry considerations:
\begin{eqnarray*} 
\begin{array}{cccc}
\cD_\theta \pars{\varPhi^h} = 0, &
\cD_\theta \pars{C^h} = 0, &
\cD_\theta \pars{V_r^h} = 0, &
V_\theta^h = 0
\end{array}
\end{eqnarray*}

The specific boundary condition at $r=1$
depend on the chemical properties of particle surface.
